---
title: "The New Way to Get Around NYC - For Who?"
author: By Natalia Iannucci and Hana Hirano
output: 
  html_document:
    code_folding: hide
    theme: journal
    df_print: paged
---
Date: `r format(Sys.Date(), "%B %e, %Y")`  
GitHub^[https://github.com/niannucci/MP4-querying] 

![Source: Curbed New York](https://cdn.vox-cdn.com/thumbor/abPMvQOI6AYVBuDmggKP7FzMppE=/0x0:3464x2309/920x613/filters:focal(1455x878:2009x1432)/cdn.vox-cdn.com/uploads/chorus_image/image/61954281/citibike.0.jpg)

Ride sharing has been increasingly popular in recent years; but what about in metropolitan cities like New York City with constant heavy traffic? Or environmentalists looking to reduce their carbon footprint? Or, maybe you just want to get a little excersize in on your way to work. No matter your motivation, Citi Bike is a bike share system that has become popular as of late. [Citi Bikes](https://www.citibikenyc.com/about) are specifically labeled bikes locked into hundreds of stations around New York City; they can be returned to any station, and are available at all times. A corresponding app is used to unlock a bike, and pay for your ride. The distance of each ride is recorded, as well as which station it was taken from and returned to and the age of the user, allowing us to examine factors that are involved in the biking world.

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(ggthemes)
library(RMySQL)
library(DBI)
db <- dbConnect(MySQL(), 
                host = "scidb.smith.edu", 
                user = "mth292", 
                password = "RememberPi", 
                dbname = "citibike")
knitr::opts_chunk$set(connection = db, max.print = 20)
```

While the Citi Bike station with the longest average duration also has the youngest average age of its users, the other stations quickly taper off and show no significant difference in average ride duration based on age.
```{r message = FALSE, warning = FALSE}
women <- dbGetQuery(db, 
"SELECT gender, AVG(t.duration / 60) AS avg_duration_in_min, AVG(2019 - t.birth_year) AS avg_age, name
FROM trips AS t
LEFT JOIN station_months 
	ON station_months.station_id = t.start_station_id
WHERE gender = 2
	AND start_time LIKE '2017-01%'
GROUP BY name
HAVING avg_age < 100
ORDER BY avg_duration_in_min DESC
LIMIT 0, 20"
)
```

```{r message = FALSE, warning = FALSE}
ggplot(women, aes(x = avg_age, y = avg_duration_in_min)) +
  ggtitle("Top 20 Longest Average Station Rides for Women") +
  geom_point(alpha = 0.8) +
  xlab("Average Age of Citi Bike User") +
  ylab("Average Ride Duration (min)") +
  theme_solarized() +
  geom_smooth()

women
```

Similarly to women, the Citi Bike station with the longest average duration also has the youngest average age of its male users, and also show no significant difference in biking duration based on the factor of age.
```{r message = FALSE, warning = FALSE}
men <- dbGetQuery(db, 
"SELECT gender, AVG(t.duration / 60) AS avg_duration_in_min, AVG(2019 - t.birth_year) AS avg_age, name
FROM trips AS t
LEFT JOIN station_months 
	ON station_months.station_id = t.start_station_id
WHERE gender = 1
	AND start_time LIKE '2017-01%'
GROUP BY name
HAVING avg_age < 100
ORDER BY avg_duration_in_min DESC
LIMIT 0, 20"
)
```

```{r message = FALSE, warning = FALSE}
ggplot(men, aes(x = avg_age, y = avg_duration_in_min)) +
  ggtitle("Top 20 Longest Average Station Rides for Men") +
  geom_point(alpha = 0.8) +
  xlab("Average Age of Citi Bike User") +
  ylab("Average Ride Duration (min)") +
  theme_solarized() +
  geom_smooth()

men
```

We also studied how people's age would affect the length of the time they use citibike. Our hypothesis was that as the person's age increases, the duration of the ride will decrease considering the physical strength. However, surprisingly, the data shows that there is almost no relationship between the age and the duration of the ride.
```{r message = FALSE, warning = FALSE}
duration_age <- dbGetQuery(db, 
"SELECT AVG(t.duration / 60) AS avg_duration_in_min, (2019 - t.birth_year) AS age
FROM trips AS t
WHERE t.birth_year > 1919
GROUP BY age
ORDER BY age"
)
```

```{r message = FALSE, warning = FALSE}
ggplot(duration_age, aes(x = age, y = avg_duration_in_min)) +
  geom_point(alpha = 0.8) +
  ggtitle("The Relationship Between Age and The Duration of Their Ride") +
  xlab("Age") +
  ylab("Average Duration of The Ride (min)") +
  theme_solarized()
```

This is an interesting trend. We expected that younger people would use the shared bike service for longer period of time since generally speaking, they should be physically strong. And we were able to see that the relationship between the age and the average duration of the ride seems to have a downward slope for the interval age 18 to around 30, meaning younger people ride londer than older people. Yet looking at the entire trend curve, for those of age 18 to 70, the trend line is almost horizontal. They seem to use the bike for approzimately 13 minutes. This horizontal trend was unexpected. Additionally, for people who are 75 or older, there seem to be two diverging trend lines, a very steep upward slope and a steep downward slope. The downward slope tells that as people age, they tend to use the shared bike servide for a shorter period of time. This makes sense as it is quite dangerous for elderly to ride bikes. Yet the upward sloping trend line was a surprise to us. This indicates that some elder people will ride shared bike for longer duration and as older they get, they use the service longer. 

> Word count: `r wordcountaddin::word_count()`